---
title: 'Intro to R - HW #2'
author: "Ted Lyman"
date: "October 16, 2015"
output: html_document
---

#####HW #2 -- _with ggplot, dplyr & troubleshooting_

###### -- please note that this is written as a notebook where I have tried differnt code and used the results to explore answers and then build upon each try.  

###### _-- Also, please note that lines are not wrapping properly nor are paragrphs being created as they are coded in the rMarkdown code written here -  something about the way the HTML code is being created through the "knitting' process"_

##### Resetting R's memory
```{r echo=TRUE}
rm(list = ls())
```


##### Load libraries for the homework
```{r echo=TRUE}
library(dplyr)
library(gapminder)
library(ggplot2)
library(tidyr)
library(formatR)
```

#### Plot 1 - create graph of the gdpPercap (x) compared to life expectancy (Y) of each continent

##### Convert df to tbl and group by vaules - because this is what we did in class
```{r echo=TRUE}
lifeExp_gdp_df <-
  gapminder %>%
  group_by(continent, lifeExp, gdpPercap)
```

##### Then plot -- like we did in class
```{r echo=TRUE}
ggplot(data = lifeExp_gdp_df, aes(x = gdpPercap, y = lifeExp)) + 
  geom_point(aes(shape = continent)) + 
  geom_point(aes(color = continent)) 
```

#### Plot 2 - Use "transformation" to linerise the data -- lm? ug... ahh a hint on the hw page -- so a "log scale" 
###### yep... there is one in ggplot... so not the "lm" function I was trying to figure out_ 
```{r echo=TRUE}
ggplot(data = lifeExp_gdp_df, aes(x = gdpPercap, y = lifeExp)) + 
  geom_point(aes(shape = continent)) + 
  geom_point(aes(color = continent)) + scale_x_log10()
```

#### Plot 3 - Include a simple "linear fit" to the transformed area
###### googled "linear fit" to get "regression line" and then got -- geom_smooth(method=lm)
```{r echo=TRUE}
ggplot(data = lifeExp_gdp_df, aes(x = gdpPercap, y = lifeExp)) + 
  geom_point(aes(shape = continent)) + 
  geom_point(aes(color = continent)) + scale_x_log10() + 
  geom_smooth(method=lm)
```

#### Plot 4 - plot the density functions of the life expectency for each continent
##### _code below **will** caouse R to hang_
ggplot(data = gapminder, aes(x = continent, y = , color = continent )) + 
  geom_area(stat="bin") + 
  facet_grid(lifeExp ~ continent)

###### _Eventually leading to this error message mess...   What in the code above brought on pages and pages of the text below?   Posting just a part of it...   hmmm....  unexpected variables...  Further more... why isn't this text and the code above wrapping to new lines as I am telling it to do??_  

...can result in unexpected behavior and will not be allowed in a future version of ggplot2.
If you want y to represent counts of cases, use stat="bin" and don't map a variable to y.
If you want y to represent values in the data, use stat="identity".
See ?geom_bar for examples. (Defunct; last used in version 0.9.2)
Error : Mapping a variable to y and also using stat="bin".
With stat="bin", it will attempt to set the y value to the count of cases in each group.
This can result in unexpected behavior and will not be allowed in a future version of ggplot2.
If you want y to represent counts of cases, use stat="bin" and don't map a #variable to y.
If you want y to represent values in the data, use stat="identity".
See ?geom_bar for examples. (Defunct; last used in version 0.9.2)
Error : Mapping a variable to y and also using stat="bin".
With stat="bin", it will attempt to set the y value to the count of cases #in each group.
This can result in unexpected behavior and will not be allowed in a future #version of ggplot2.
If you want y to represent counts of cases, use stat="bin" and don't map a variable to y.
If you want y to represent values in the data, use stat="identity".
See ?geom_bar for examples. (Defunct; last used in version 0.9.2)



#### _ok - so no "y"?? ... nope y=lifeExp_
##### well - something popped up but then disappeared..  ??
```{r echo=TRUE}
ggplot(data = gapminder, aes(x = continent, y = lifeExp, color = continent )) +  
geom_boxplot() 
```

#### _Somethings very not right... no plots being displayed anymore..._

##### reset Rstudio...  trying something new
```{r echo=TRUE}
gdp_df <-
  gapminder %>%
  group_by(continent, lifeExp)

ggplot(data =gdp_df, aes(x = gdpPercap, y = lifeExp)) + 
  geom_point(aes(shape = continent)) + 
  geom_point(aes(color = continent)) 
```

###### ok... this looks like it works - refining

```{r echo=TRUE}
ggplot(data = gdp_df, aes(x = continent, y = lifeExp, color = continent)) + 
  geom_boxplot() +
  facet_grid(year~continent)
```

###### wrong plot type but the grid looks promising

```{r echo=TRUE}
ggplot(data = gdp_df, aes(x = continent, y = lifeExp, color = continent)) + 
  geom_boxplot() +
  facet_wrap(year~continent)
```  

###### nope...

```{r echo=TRUE}
ggplot(data = gdp_df, aes(x = continent, y = lifeExp, color = continent)) + 
  geom_area() +  facet_wrap(~year)
```

###### nope...  

```{r echo=TRUE}
ggplot(data = gdp_df, aes(x = lifeExp, y = , color = continent, fill = continent)) + 
  geom_density(color = 'black', alpha = 0.5) +  facet_wrap(~year)
```  

###### got it with Nelson's help missing the 'fill' part as well as the density params  

#### Plot 5 - fix the wording...  create the plot first of course.... well - it's a boxplot based on the above trials...  

```{r echo=TRUE}
ggplot(data = gdp_df, aes(x = continent, y = lifeExp, color = continent)) + 
  geom_boxplot() +  facet_wrap(~year)
```

###### (ummm... just widen the screen... mine looks just fine at full screen..??)
###### ggplot cheat sheet says something about using scale function when looking at the "Labels" tab... don't see how to do this..)

```{r echo=TRUE}
ggplot(data = gdp_df, aes(x = continent, y = lifeExp, color = continent)) + 
  geom_boxplot() + 
  facet_wrap(~year) + 
  theme_minimal()
```

###### like the minimal theme - but it didn't fix the text...  looks like it could be "scale_fill_discrete" 

```{r echo=TRUE}
ggplot(data = gdp_df, aes(x = continent, y = lifeExp, color = continent)) + 
  geom_boxplot() + 
  facet_wrap(~year) + 
  theme_minimal() + 
  scale_fill_discrete()
```

###### giving up... googling it... yep - This fixes it

```{r echo=TRUE}
ggplot(data = gdp_df, aes(x = continent, y = lifeExp, color = continent)) + 
  geom_boxplot() +  
  facet_wrap(~year) + 
  theme(axis.text.x=element_text(angle=50, size=10, vjust=0.5)) + 
  scale_fill_discrete()
```

###### Plot 6 - Use dplyr to calc summary stats  (copying text from dplr class and working from that)
###### _(make sure "dplyr" reloads when restarting RStudio)_

```{r echo=TRUE}
select(gapminder)
tbl_df(gapminder)
group_by(gapminder)
```

```{r echo=TRUE}
ggplot(data = gapminder, aes(x = lifeExp, y = , fill = continent)) +
  geom_density(color = "black", alpha = 0.5)
```

#### Plot 7 - Life expectancy in Asia

```{r echo=TRUE}
a_data <- filter(gapminder, continent == "Asia")
ggplot(data = a_data, aes(x = lifeExp, y = , color = continent, 
                              fill = continent)) +
  scale_fill_manual(values = c("green")) + 
  geom_density(color = "black", alpha = 0.5) + 
  geom_vline(aes(xintercept = mean(lifeExp))) +
  theme_minimal() + 
  ggtitle("Life expectancy in Asia")
```

#### Dplyr Plot 8a - Create a dataframe of the mean life expectancies for each continent

```{r echo=TRUE}
mean_lifeExp <- 
  gapminder %>%
  group_by(continent) %>%
  summarize(avg = mean(lifeExp))
```
###### _Summarize takes big thing and makes it smaller_

#### Dplyr Plot 8b - Plot the densities of the life expectancies for each continent and draw a line to mark the mean for each continent

```{r echo=TRUE}
ggplot(data = gapminder, aes(x = lifeExp, y = , colour = continent, 
                             fill = continent)) +
  facet_wrap(~continent) + 
  geom_density(color = "black", alpha = 0.6) + 
  geom_vline(data = mean_lifeExp, aes(xintercept = avg))
```

##### Troubleshooting -- Finding mistakes in listed code
##### Example code is displayed only - and not run, as..., um..., duh... it generates an error  

```{r echo=TRUE}
library(ggplot2)
hw_gapminder <- read.csv('/Users/lyman/Dropbox (Byrnes Lab)/R_projects/Intro to R_Fall 2015/hw_gapminder.csv')
```

##### 1)
> mean-lifeExp <- mean(hw_gapminder$lifeExp**e**)

###### Ha!  The first example - the extra 'e' at the end of 'lifeExpe' was autofixed when I typed it in.   Nice!   

```{r echo=TRUE}
mean_lifeExp <- mean(hw_gapminder$lifeExp)
```

##### 2)
> small-set <- hw_gapminder[c(1, 2, 3, 4, 1300:1304), ('country', 'continent', 'year')]

###### Ha again!  auto formatting has already picked out the lonely commas showing where the error is...  which looks like a 'cat' problem  

```{r echo=TRUE}
small_set <- hw_gapminder[c(1, 2, 3, 4, 1300:1304), c('country', 'continent', 'year')]
```
###### Yup - concatenate issue  

##### 3)
> mean-gdp <- mean(hw_gapminder$gdpPercap, na.rm = TRUE)

###### Running this returns a "NA_real_ value error.  From class - you have check data and in this case NA is not a number you can get the mean of --- thanks also to Marc for pointing this out _AND_ giving me the 'na.rm = TRUE' answer  

```{r echo=TRUE}
mean_gdp <- mean(hw_gapminder$gdpPercap, na.rm = TRUE)
```
##### 4)
> max-country <- hw-gapminder$country [which(hw-gapminder$lifeExp = max(hw-gapminder$lifeExp))]

###### Equal sign error -- need the second '=' to make '==' as 'equal to'  

```{r echo=TRUE}
max_country <- hw_gapminder$country[which(hw_gapminder$lifeExp == max(hw_gapminder$lifeExp))]
```

